National Longitudinal Evidence for Growth in Subjective Well-being from Spiritual Beliefs

### 2021-03-11 Manuscript_Bayesian.Rmd
updated manuscript

A second challenge is to disentangle concepts of spirituality from concepts of religion. The term "spirituality" was first used to describe the religious practices of ascetics and monks, however, more recently, the term has taken on a broader family of meanings that some of which apply to non-religious contexts [@Swinton2001-vr]. Notably, spirituality and health research has not converged on a unique set of meanings. Although it is commonplace across most areas of psychological science for researchers to investigate different operationalisations of focal concepts (think of memory research, or emotions research), the psychology of spirituality has mostly been conducted with cross-sectional North American samples among participants who affiliate with traditional religion [@Ano2005-hx; @Hackney2003-rs; @Sawatzky2005-rw; @Smith2003-re; @Visser2010-kq; @Yonker2012-zg]. Whether the psychology of spirituality and health has repeated or extended the psychology of religion and health remains unclear. Put simply, cnceptual overlap between spirituality constructs and religion constructs creates a second line of terminological confusion that interacts with terminological confusion arising from overlap between spirituality constructs and well-being constructs. Indeed, researchers who have operationalised spirituality constructs as distinct from religion constructs have observed a negative association between spirituality and subjective well-being [@King2013-cg]. Thus to infer how spirituality affects subjective well-being, it is not sufficient to disentange spirituality constructs from tautological well-being constructs, it is also important to disentangle the concept of spirituality from the concept of religion, and to carefully investigate the mechanisms of well-being in populations who overlap, partially overlap, and differ in these dimensions.

A third challenge is to investigate the relationship between spirituality and subjective well-being in nationally diverse samples over individual life-spans. To assess the mechanisms of well-being production requires observing change within people who differ in theoregically relevant ways and assessing how their lives turn out [@Garssen2016-km]. However, previous investigations of spirituality and well-being, including high-quality reviews and meta-analyses, have nearly exclusively relied on cross-sectional samples [@Ano2005-hx; @Hackney2003-rs; @Sawatzky2005-rw; @Smith2003-re; @Visser2010-kq; @Yonker2012-zg]. Although researching the key questions in spirituality psychological health research requires national longitudinal investigations, unfortunately the relevant datasets are rare.

@@

Because missing values can result in biased estimates [@Blackwell2017-oq], we multiply imputed 10 datasets using the AmeliaII package in R[@Honaker2011-yu].We conducted a parallel analysis averaging over the multiply imputed datasets, without any substantive difference to inferences.

@@

The analysis was performed using R version 4.0.2 (2020-06-22). The Platform was x86_64-apple-darwin17.0 (64-bit) Running under: macOS Catalina 10.15.6 We are greatful to the contributors and mantainers of the following packages: R [Version 4.0.3; @R-base] and the R-packages *Amelia* [Version 1.7.6; @R-Amelia], *citr* [Version 0.3.2; @R-citr], *dplyr* [Version 1.0.2; @R-dplyr], *equatiomatic* [Version 0.1.0.9000; @R-equatiomatic], *forcats* [Version 0.5.0; @R-forcats], *ggeffects* [Version 1.0.0; @R-ggeffects], *gghighlight* [Version 0.3.0; @R-gghighlight], *ggplot2* [Version 3.3.2; @R-ggplot2], *ggsci* [Version 2.9; @R-ggsci], *lme4* [Version 1.1.25; @R-lme4], *Matrix* [Version 1.2.18; @R-Matrix], *papaja* [Version 0.1.0.9997; @R-papaja], *parameters* [Version 0.9.0.1; @R-parameters], *patchwork* [Version 1.1.0; @R-patchwork], *prettycode* [@R-prettycode], *purrr* [Version 0.3.4; @R-purrr], *Rcpp* [Version 1.0.5; @R-Rcpp_a; @R-Rcpp_b], *readr* [Version 1.4.0; @R-readr], *report* [Version 0.2.0; @R-report], *see* [Version 0.6.0.1; @R-see], *sjPlot* [Version 2.8.6; @R-sjPlot], *stringr* [Version 1.4.0; @R-stringr], *styler* [Version 1.3.2; @R-styler], *table1* [Version 1.2.1; @R-table1], *texreg* [Version 1.37.5; @R-texreg], *tibble* [Version 3.0.4; @R-tibble], *tidyr* [Version 1.1.2; @R-tidyr], and *tidyverse* [Version 1.3.0; @R-tidyverse]

@@

All code for this analysis and a full report of the missing data imputation can be found at [https://github.com/go-bayes/nzavs_spiritbeliefWellbeing](https://github.com/go-bayes/nzavs_spiritbeliefWellbeing){.uri}. An html document decribing the same analysis is located at: <https://osf.io/cu2gr/>.

@@

Focusing on life satisfaction, people who believed in Spirit, either with a belief in God ($\hat{y}$ = 5.18 \| [5.13, 5.24]) or without a belief in God (5.18 \| [5.12, 5.24]) tended to be lower in life satsifaction than disbelievers in both a God and a Spirit (5.27 \| [5.20, 5.33]) and disbelievers in a spirit who also believed in a God (5.26 \| [5.15, 5.36]) (see Figure 3)

@@

Because These inferred growth rates for personal well being and life satisfaction are evident from repeated measures across 10 annual waves of nationally diverse data collection, we are confident that in New Zealand, spiritual beliefs predict strong increases in personal well-being and life satisfaction

>>>>

A second challenge is to disentangle concepts of spirituality from concepts of religion. The term "spirituality" was first used to describe the religious practices of ascetics and monks, however, more recently, the term has taken on a broader family of meanings that some of which apply to non-religious contexts [@Swinton2001-vr]. Notably, spirituality and health research has not converged on a unique set of meanings. Although it is commonplace across most areas of psychological science for researchers to investigate different operationalisations of focal concepts (think of memory research, or emotions research), the psychology of spirituality has mostly been conducted with cross-sectional North American samples among participants who affiliate with traditional religion [@Ano2005-hx; @Hackney2003-rs; @Sawatzky2005-rw; @Smith2003-re; @Visser2010-kq; @Yonker2012-zg; @holt2003spiritual]. Whether the psychology of spirituality and health has repeated or extended the psychology of religion and health remains unclear. Put simply, conceptual overlap between spirituality constructs and religion constructs creates a second line of terminological confusion that interacts with terminological confusion arising from overlap between spirituality constructs and well-being constructs. Indeed, researchers who have operationalised spirituality constructs as distinct from religion constructs have observed a negative association between spirituality and subjective well-being [@King2013-cg]. Thus to infer how spirituality affects subjective well-being, it is not sufficient to disentangle spirituality constructs from tautological well-being constructs, it is also important to disentangle the concept of spirituality from the concept of religion, and to carefully investigate the mechanisms of well-being in populations who overlap, partially overlap, and differ in these dimensions.

A third challenge is to investigate the relationship between spirituality and subjective well-being in nationally diverse samples over individual life-spans. To assess the mechanisms of well-being production requires observing change within people who differ in theoretically relevant ways and assessing how their lives turn out [@Garssen2016-km]. However, previous investigations of spirituality and well-being, including high-quality reviews and meta-analyses, have nearly exclusively relied on cross-sectional samples [@Ano2005-hx; @Hackney2003-rs; @Sawatzky2005-rw; @Smith2003-re; @Visser2010-kq; @Yonker2012-zg; @januu2020associations], and sampled from small sub-populations such as older adults [@lawler2009role], urban African American women [@holt2003spiritual], rural Kenyan men [@goodman2020mental], and Ghanaian parents of children with special needs [@dey2019spirituality]. Such limitations have been evident over 20 years (see for example: [@thoresen1999spirituality]. However, until recently, national-scale longitudinal time-series data required to investigate the long term effects of spirituality have been lacking.

@@

Because missing values can result in biased estimates [@Blackwell2017-oq], we multiply imputed 10 datasets using the Amelia II package in R[@Honaker2011-yu].We conducted a parallel analysis averaging over the multiply imputed datasets, without any substantive difference to inferences.

@@

The analysis was performed using R version 4.0.2 (2020-06-22). The Platform was x86_64-apple-darwin17.0 (64-bit) Running under: macOS Catalina 10.15.6 We are grateful to the contributors and maintainers of the following packages: R [Version 4.0.3; @R-base] and the R-packages *Amelia* [Version 1.7.6; @R-Amelia], *citr* [Version 0.3.2; @R-citr], *dplyr* [Version 1.0.2; @R-dplyr], *equatiomatic* [Version 0.1.0.9000; @R-equatiomatic], *forcats* [Version 0.5.0; @R-forcats], *ggeffects* [Version 1.0.0; @R-ggeffects], *gghighlight* [Version 0.3.0; @R-gghighlight], *ggplot2* [Version 3.3.2; @R-ggplot2], *ggsci* [Version 2.9; @R-ggsci], *lme4* [Version 1.1.25; @R-lme4], *Matrix* [Version 1.2.18; @R-Matrix], *papaja* [Version 0.1.0.9997; @R-papaja], *parameters* [Version 0.9.0.1; @R-parameters], *patchwork* [Version 1.1.0; @R-patchwork], *prettycode* [@R-prettycode], *purrr* [Version 0.3.4; @R-purrr], *Rcpp* [Version 1.0.5; @R-Rcpp_a; @R-Rcpp_b], *readr* [Version 1.4.0; @R-readr], *report* [Version 0.2.0; @R-report], *see* [Version 0.6.0.1; @R-see], *sjPlot* [Version 2.8.6; @R-sjPlot], *stringr* [Version 1.4.0; @R-stringr], *styler* [Version 1.3.2; @R-styler], *table1* [Version 1.2.1; @R-table1], *texreg* [Version 1.37.5; @R-texreg], *tibble* [Version 3.0.4; @R-tibble], *tidyr* [Version 1.1.2; @R-tidyr], and *tidyverse* [Version 1.3.0; @R-tidyverse]

@@

All code for this analysis and a full report of the missing data imputation can be found at [https://github.com/go-bayes/nzavs_spiritbeliefWellbeing](https://github.com/go-bayes/nzavs_spiritbeliefWellbeing){.uri}. An html document describing the same analysis is located at: <https://osf.io/cu2gr/>.

@@

Focusing on life satisfaction, people who believed in Spirit, either with a belief in God ($\hat{y}$ = 5.18 \| [5.13, 5.24]) or without a belief in God (5.18 \| [5.12, 5.24]) tended to be lower in life satisfaction than disbelievers in both a God and a Spirit (5.27 \| [5.20, 5.33]) and disbelievers in a spirit who also believed in a God (5.26 \| [5.15, 5.36]) (see Figure 3)

@@

Because these inferred growth rates for personal well being and life satisfaction are evident from repeated measures across 10 annual waves of nationally diverse data collection, we are confident that in New Zealand, spiritual beliefs predict strong increases in personal well-being and life satisfaction

### 2021-03-11 Manuscript_Bayesian.tex
updated manuscript

A second challenge is to disentangle concepts of spirituality from concepts of religion. The term ``spirituality'' was first used to describe the religious practices of ascetics and monks, however, more recently, the term has taken on a broader family of meanings that some of which apply to non-religious contexts (Swinton, 2001). Notably, spirituality and health research has not converged on a unique set of meanings. Although it is commonplace across most areas of psychological science for researchers to investigate different operationalisations of focal concepts (think of memory research, or emotions research), the psychology of spirituality has mostly been conducted with cross-sectional North American samples among participants who affiliate with traditional religion (Ano \& Vasconcelles, 2005; Hackney \& Sanders, 2003; Sawatzky, Ratner, \& Chiu, 2005; Smith, McCullough, \& Poll, 2003; Visser, Garssen, \& Vingerhoets, 2010; Yonker, Schnabelrauch, \& Dehaan, 2012). Whether the psychology of spirituality and health has repeated or extended the psychology of religion and health remains unclear. Put simply, cnceptual overlap between spirituality constructs and religion constructs creates a second line of terminological confusion that interacts with terminological confusion arising from overlap between spirituality constructs and well-being constructs. Indeed, researchers who have operationalised spirituality constructs as distinct from religion constructs have observed a negative association between spirituality and subjective well-being (King et al., 2013). Thus to infer how spirituality affects subjective well-being, it is not sufficient to disentange spirituality constructs from tautological well-being constructs, it is also important to disentangle the concept of spirituality from the concept of religion, and to carefully investigate the mechanisms of well-being in populations who overlap, partially overlap, and differ in these dimensions.

A third challenge is to investigate the relationship between spirituality and subjective well-being in nationally diverse samples over individual life-spans. To assess the mechanisms of well-being production requires observing change within people who differ in theoregically relevant ways and assessing how their lives turn out (Garssen \& Visser, 2016). However, previous investigations of spirituality and well-being, including high-quality reviews and meta-analyses, have nearly exclusively relied on cross-sectional samples (Ano \& Vasconcelles, 2005; Hackney \& Sanders, 2003; Sawatzky, Ratner, \& Chiu, 2005; Smith, McCullough, \& Poll, 2003; Visser, Garssen, \& Vingerhoets, 2010; Yonker, Schnabelrauch, \& Dehaan, 2012). Although researching the key questions in spirituality psychological health research requires national longitudinal investigations, unfortunately the relevant datasets are rare.

@@

Because missing values can result in biased estimates (Blackwell, Honaker, \& King, 2017), we multiply imputed 10 datasets using the AmeliaII package in R(Honaker, King, Blackwell, \& Others, 2011).We conducted a parallel analysis averaging over the multiply imputed datasets, without any substantive difference to inferences.

@@

The analysis was performed using R version 4.0.2 (2020-06-22). The Platform was x86\_64-apple-darwin17.0 (64-bit) Running under: macOS Catalina 10.15.6 We are greatful to the contributors and mantainers of the following packages: R {[}Version 4.0.3; R Core Team (2020){]} and the R-packages \emph{Amelia} {[}Version 1.7.6; Honaker, King, and Blackwell (2011){]}, \emph{citr} {[}Version 0.3.2; Aust (2019){]}, \emph{dplyr} {[}Version 1.0.2; Wickham, François, Henry, and Müller (2020){]}, \emph{equatiomatic} {[}Version 0.1.0.9000; Anderson and Heiss (2020){]}, \emph{forcats} {[}Version 0.5.0; Wickham (2020a){]}, \emph{ggeffects} {[}Version 1.0.0; Lüdecke (2018){]}, \emph{gghighlight} {[}Version 0.3.0; Yutani (2020){]}, \emph{ggplot2} {[}Version 3.3.2; Wickham (2016){]}, \emph{ggsci} {[}Version 2.9; Xiao (2018){]}, \emph{lme4} {[}Version 1.1.25; Bates, Mächler, Bolker, and Walker (2015){]}, \emph{Matrix} {[}Version 1.2.18; Bates and Maechler (2019){]}, \emph{papaja} {[}Version 0.1.0.9997; Aust and Barth (2020){]}, \emph{parameters} {[}Version 0.9.0.1; Lüdecke, Ben-Shachar, Patil, and Makowski (2020){]}, \emph{patchwork} {[}Version 1.1.0; Pedersen (2020){]}, \emph{prettycode} (Csárdi, 2019), \emph{purrr} {[}Version 0.3.4; Henry and Wickham (2020){]}, \emph{Rcpp} {[}Version 1.0.5; Eddelbuettel and François (2011); Eddelbuettel and Balamuta (2017){]}, \emph{readr} {[}Version 1.4.0; Wickham, Hester, and Francois (2018){]}, \emph{report} {[}Version 0.2.0; Makowski, Dominique, Lüdecke, and Daniel (2019){]}, \emph{see} {[}Version 0.6.0.1; Lüdecke, Ben-Shachar, Waggoner, and Makowski (2020){]}, \emph{sjPlot} {[}Version 2.8.6; Lüdecke (2020){]}, \emph{stringr} {[}Version 1.4.0; Wickham (2019){]}, \emph{styler} {[}Version 1.3.2; Müller and Walthert (2020){]}, \emph{table1} {[}Version 1.2.1; Rich (2020){]}, \emph{texreg} {[}Version 1.37.5; Leifeld (2013){]}, \emph{tibble} {[}Version 3.0.4; Müller and Wickham (2020){]}, \emph{tidyr} {[}Version 1.1.2; Wickham (2020b){]}, and \emph{tidyverse} {[}Version 1.3.0; Wickham et al. (2019){]}

@@

All code for this analysis and a full report of the missing data imputation can be found at \url{https://github.com/go-bayes/nzavs_spiritbeliefWellbeing}. An html document decribing the same analysis is located at: \url{https://osf.io/cu2gr/}.

@@

Focusing on life satisfaction, people who believed in Spirit, either with a belief in God (\(\hat{y}\) = 5.18 \textbar{} {[}5.13, 5.24{]}) or without a belief in God (5.18 \textbar{} {[}5.12, 5.24{]}) tended to be lower in life satsifaction than disbelievers in both a God and a Spirit (5.27 \textbar{} {[}5.20, 5.33{]}) and disbelievers in a spirit who also believed in a God (5.26 \textbar{} {[}5.15, 5.36{]}) (see Figure 3)

@@

@@

@@

@@

@@

@@

>>>>

A second challenge is to disentangle concepts of spirituality from concepts of religion. The term ``spirituality'' was first used to describe the religious practices of ascetics and monks, however, more recently, the term has taken on a broader family of meanings that some of which apply to non-religious contexts (Swinton, 2001). Notably, spirituality and health research has not converged on a unique set of meanings. Although it is commonplace across most areas of psychological science for researchers to investigate different operationalisations of focal concepts (think of memory research, or emotions research), the psychology of spirituality has mostly been conducted with cross-sectional North American samples among participants who affiliate with traditional religion (Ano \& Vasconcelles, 2005; Hackney \& Sanders, 2003; Holt, Clark, Kreuter, \& Rubio, 2003; Sawatzky, Ratner, \& Chiu, 2005; Smith, McCullough, \& Poll, 2003; Visser, Garssen, \& Vingerhoets, 2010; Yonker, Schnabelrauch, \& Dehaan, 2012). Whether the psychology of spirituality and health has repeated or extended the psychology of religion and health remains unclear. Put simply, conceptual overlap between spirituality constructs and religion constructs creates a second line of terminological confusion that interacts with terminological confusion arising from overlap between spirituality constructs and well-being constructs. Indeed, researchers who have operationalised spirituality constructs as distinct from religion constructs have observed a negative association between spirituality and subjective well-being (King et al., 2013). Thus to infer how spirituality affects subjective well-being, it is not sufficient to disentangle spirituality constructs from tautological well-being constructs, it is also important to disentangle the concept of spirituality from the concept of religion, and to carefully investigate the mechanisms of well-being in populations who overlap, partially overlap, and differ in these dimensions.

A third challenge is to investigate the relationship between spirituality and subjective well-being in nationally diverse samples over individual life-spans. To assess the mechanisms of well-being production requires observing change within people who differ in theoretically relevant ways and assessing how their lives turn out (Garssen \& Visser, 2016). However, previous investigations of spirituality and well-being, including high-quality reviews and meta-analyses, have nearly exclusively relied on cross-sectional samples (Ano \& Vasconcelles, 2005; Hackney \& Sanders, 2003; Jan, Malinakova, Kosarkova, \& Tavel, 2020; Sawatzky, Ratner, \& Chiu, 2005; Smith, McCullough, \& Poll, 2003; Visser, Garssen, \& Vingerhoets, 2010; Yonker, Schnabelrauch, \& Dehaan, 2012), and sampled from small sub-populations such as older adults (Lawler-Row \& Elliott, 2009), urban African American women (Holt, Clark, Kreuter, \& Rubio, 2003), rural Kenyan men (Goodman, Gitari, Keiser, Elliott, \& Seidel, 2020), and Ghanaian parents of children with special needs (Dey, Amponsah, \& Wiafe-Akenteng, 2019). Such limitations have been evident over 20 years (see for example: (Thoresen, 1999). However, until recently, national-scale longitudinal time-series data required to investigate the long term effects of spirituality have been lacking.

@@

Because missing values can result in biased estimates (Blackwell, Honaker, \& King, 2017), we multiply imputed 10 datasets using the Amelia II package in R(Honaker, King, Blackwell, \& Others, 2011).We conducted a parallel analysis averaging over the multiply imputed datasets, without any substantive difference to inferences.

@@

The analysis was performed using R version 4.0.2 (2020-06-22). The Platform was x86\_64-apple-darwin17.0 (64-bit) Running under: macOS Catalina 10.15.6 We are grateful to the contributors and maintainers of the following packages: R {[}Version 4.0.3; R Core Team (2020){]} and the R-packages \emph{Amelia} {[}Version 1.7.6; Honaker, King, and Blackwell (2011){]}, \emph{citr} {[}Version 0.3.2; Aust (2019){]}, \emph{dplyr} {[}Version 1.0.2; Wickham, François, Henry, and Müller (2020){]}, \emph{equatiomatic} {[}Version 0.1.0.9000; Anderson and Heiss (2020){]}, \emph{forcats} {[}Version 0.5.0; Wickham (2020a){]}, \emph{ggeffects} {[}Version 1.0.0; Lüdecke (2018){]}, \emph{gghighlight} {[}Version 0.3.0; Yutani (2020){]}, \emph{ggplot2} {[}Version 3.3.2; Wickham (2016){]}, \emph{ggsci} {[}Version 2.9; Xiao (2018){]}, \emph{lme4} {[}Version 1.1.25; Bates, Mächler, Bolker, and Walker (2015){]}, \emph{Matrix} {[}Version 1.2.18; Bates and Maechler (2019){]}, \emph{papaja} {[}Version 0.1.0.9997; Aust and Barth (2020){]}, \emph{parameters} {[}Version 0.9.0.1; Lüdecke, Ben-Shachar, Patil, and Makowski (2020){]}, \emph{patchwork} {[}Version 1.1.0; Pedersen (2020){]}, \emph{prettycode} (Csárdi, 2019), \emph{purrr} {[}Version 0.3.4; Henry and Wickham (2020){]}, \emph{Rcpp} {[}Version 1.0.5; Eddelbuettel and François (2011); Eddelbuettel and Balamuta (2017){]}, \emph{readr} {[}Version 1.4.0; Wickham, Hester, and Francois (2018){]}, \emph{report} {[}Version 0.2.0; Makowski, Dominique, Lüdecke, and Daniel (2019){]}, \emph{see} {[}Version 0.6.0.1; Lüdecke, Ben-Shachar, Waggoner, and Makowski (2020){]}, \emph{sjPlot} {[}Version 2.8.6; Lüdecke (2020){]}, \emph{stringr} {[}Version 1.4.0; Wickham (2019){]}, \emph{styler} {[}Version 1.3.2; Müller and Walthert (2020){]}, \emph{table1} {[}Version 1.2.1; Rich (2020){]}, \emph{texreg} {[}Version 1.37.5; Leifeld (2013){]}, \emph{tibble} {[}Version 3.0.4; Müller and Wickham (2020){]}, \emph{tidyr} {[}Version 1.1.2; Wickham (2020b){]}, and \emph{tidyverse} {[}Version 1.3.0; Wickham et al. (2019){]}

@@

All code for this analysis and a full report of the missing data imputation can be found at \url{https://github.com/go-bayes/nzavs_spiritbeliefWellbeing}. An html document describing the same analysis is located at: \url{https://osf.io/cu2gr/}.

@@

Focusing on life satisfaction, people who believed in Spirit, either with a belief in God (\(\hat{y}\) = 5.18 \textbar{} {[}5.13, 5.24{]}) or without a belief in God (5.18 \textbar{} {[}5.12, 5.24{]}) tended to be lower in life satisfaction than disbelievers in both a God and a Spirit (5.27 \textbar{} {[}5.20, 5.33{]}) and disbelievers in a spirit who also believed in a God (5.26 \textbar{} {[}5.15, 5.36{]}) (see Figure 3)

@@

\leavevmode\hypertarget{ref-dey2019spirituality}{}%
Dey, N. E. Y., Amponsah, B., \& Wiafe-Akenteng, C. B. (2019). Spirituality and subjective well-being of ghanaian parents of children with special needs: The mediating role of resilience. \emph{Journal of Health Psychology}, 1359105319873956.

@@

\leavevmode\hypertarget{ref-goodman2020mental}{}%
Goodman, M. L., Gitari, S., Keiser, P., Elliott, A., \& Seidel, S. (2020). Mental health and childhood memories among rural kenyan men: Considering the role of spirituality in life-course pathways. \emph{Journal of Health Psychology}, 1359105320944984.

@@

\leavevmode\hypertarget{ref-holt2003spiritual}{}%
Holt, C. L., Clark, E. M., Kreuter, M. W., \& Rubio, D. M. (2003). Spiritual health locus of control and breast cancer beliefs among urban african american women. \emph{Health Psychology}, \emph{22}(3), 294.

@@

\leavevmode\hypertarget{ref-januu2020associations}{}%
Jan, A., Malinakova, K., Kosarkova, A., \& Tavel, P. (2020). Associations of childhood trauma experiences with religious and spiritual struggles. \emph{Journal of Health Psychology}, 1359105320950793.

@@

\leavevmode\hypertarget{ref-lawler2009role}{}%
Lawler-Row, K. A., \& Elliott, J. (2009). The role of religious activity and spirituality in the health and well-being of older adults. \emph{Journal of Health Psychology}, \emph{14}(1), 43--52.

@@

\leavevmode\hypertarget{ref-thoresen1999spirituality}{}%
Thoresen, C. E. (1999). Spirituality and health: Is there a relationship? \emph{Journal of Health Psychology}, \emph{4}(3), 291--300.
